Finch: Sparse and Structured Array Programming with Control Flow

### 2024-04-05 sections/abstract.tex
fixes for completeness
packing

In this work, we propose a new programming language, Finch, which supports \textit{both}
flexible control flow and diverse data structures. Finch facilitates a programming model which
resolves the challenges of computing over structured arrays by combining control flow and data
structures into a common representation where they can be co-optimized. In particular,
Finch automatically specializes the control flow to the data so that performance engineers can
focus on experimenting with many algorithms. Finch supports a familiar programming language
of loops, statements, if conditions, breaks, etc., over a wide variety of array structures, such as
sparsity, run-length-encoding, symmetry, triangles, padding, or blocks. Finch reliably utilizes the key
properties of structure, such as structural zeros, repeated values, or clustered non-zeros. We show that this leads to dramatic speedups in operations such as spmv and spgemm, image processing, graph analytics, and a high-level tensor operator fusion interface.
\end{abstract}

>>>>

In this work, we propose a new programming language, Finch, which supports \textit{both} flexible control flow and diverse data structures. Finch facilitates a programming model which resolves the challenges of computing over structured arrays by combining control flow and data structures into a common representation where they can be co-optimized. Finch automatically specializes control flow to data so that performance engineers can focus on experimenting with many algorithms. Finch supports a familiar programming language of loops, statements, ifs, breaks, etc., over a wide variety of array structures, such as sparsity, run-length-encoding, symmetry, triangles, padding, or blocks. Finch reliably utilizes the key properties of structure, such as structural zeros, repeated values, or clustered non-zeros. We show that this leads to dramatic speedups in operations such as spmv and spgemm, image processing, graph analytics, and a high-level tensor operator fusion interface.
\end{abstract}
